### 2018-04-13 inst/paper/diagram_model.Rnw

\documentclass{standalone}
\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\begin{document}

\tikzstyle{decision} = [diamond, draw, fill=gray!20, text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=white!20, text width=6em, text centered, rounded corners, minimum height=3em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=white!20, node distance=3cm, minimum height=2.5em]

\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
     \node [cloud] (F) {$Shared\ Component_{i}$ ($F_{i}$)};
     \node [block, above left= 1cm and -1.5cm of F] (T1) {$Mother Age_{1i}$ ($T_{1i}$)};
     \node [block, below left= 1cm and -1.5cm of F] (T2) {$Mother Age_{2i}$ ($T_{2i}$)};
     \node [block, right of=T1, node distance=6cm] (D1) {$Difficulty_{1i}$ ($D_{1i}$)};
     \node [block, right of=T2, node distance=6cm] (D2) {$Difficulty_{2i}$ ($D_{2i}$)};
     \node [right of=D1, node distance=3cm] (E1) {$\epsilon_{1i}$};
     \node [right of=D2, node distance=3cm] (E2) {$\epsilon_{2i}$};
     \node [below right= 2.5cm and 5.3cm of F] (ee) {$ $};
     \node [below right= 2.5cm and -2.5cm of F] (eek) {$ $};
     \node [below right= 0cm and 5.8cm of F] (ee2) {$\sigma_{\epsilon_{1i}T_{2i}}$};
     \node [left of=F, node distance=3.9cm] (stt) {$\sigma_{T_{1i}T_{2i}}$};
     \node [above left= .33cm and .1cm of F] (st1f) {$\sigma_{T_{1i}F_{i}}$};
     \node [below left= .33cm and .1cm of F] (st2f) {$\sigma_{T_{2i}F_{i}}$};
     \node [above right= .29cm and .4cm of F] (sfD1) {$\gamma_{F_{i}D_{1i}} = 1$};
     \node [below right= .3cm and .4cm of F] (sfD2) {$\gamma_{F_{i}D_{2i}} = 1$};
     \node [above right= 1.5cm and -1.4cm of F] (stD1) {$\gamma_{T_{1i}D_{1i}}$};
     \node [below right= 1.55cm and -1.4cm of F] (stD2) {$\gamma_{T_{2i}D_{2i}}$};
    % Draw edges
    \draw [->] (T1) to (D1);
    \draw [->] (T2) to (D2);
    \draw [->] (F) to (D1);
    \draw [->] (F) to (D2);
    \draw [->] (E1) to (D1);
    \draw [->] (E2) to (D2);
    \path (T1) edge [<->, bend right=90, looseness=1.5] node {$ $} (T2);
    \path (T1) edge [<->, out=180,in=180] node {$ $} (F);
    \path (T2) edge [<->, out=180,in=180] node {$ $} (F);
    \draw [->] (E1) to [out=0,in=0] (ee)
    to [out=180,in=0, looseness=.35] (eek)
    to [out=180,in=270, looseness=1] (T2);
\end{tikzpicture}

\end{document}